On the number of tiles touched by a\\ line segment on a rectangular grid

### 2021-08-01 paper.tex
Quick note for the "to-consider" list
Wording
Make the three problems clearer

The question of determining how many tiles a segment of a given length touches on average will be referred to as the \emph{average-number} problem. As will be seen, solving this also answers the inverse question (segment length to touch a given number of tiles on average), which thus need not be considered separately. By symmetry, one endpoint of the segment can be assumed to lie in a fixed, reference tile. The position of this endpoint will be considered to be \emph{uniformly} distributed on the tile. The segment orientation will be have a \emph{uniform} distribution on the set of all possible directions, and is \emph{independent} of the endpoint position.

@@

The problems addressed in this paper are related to \ldots ***maybe a short paragraph here mentioning related problems: Buffon's needle, Buffon-Laplace problem. Some references: ``Buffon's Noodle Problem'', J.F.~Ramaley, 1969. ``On Laplace's Extension of the Buffon Needle Problem'', B.J.~Arnow, 1994. \\
\url{https://www.jstor.org/stable/2317945} \\
\url{https://www.jstor.org/stable/2687085}

The rest of the paper is organized as follows. \S\ref{part: fund results} presents fundamental results, which form the basis of the subsequent analysis. The direct and inverse problems are considered in \S\ref{part: max}, first for real-valued lengths, and then for integer-valued lengths on a unit square grid. The average-number problem is analyzed in \S\ref{part: ave}.

@@

\section{Maximum number of tiles: direct and inverse problems}
\label{part: max}
% ***Is it clear enough that "maximum number of tiles" can refer to the inverse problem too? (because the inverse problem is closely related to the direct one, as discussed in the introduction)
% If not, an alernative would be a more explicit title like: "Maximum number of tiles fo a given length. Conditions on length for a given number of tiles"

The direct and inverse problems for the maximum number of tiles are considered in this section. The general case for rectangular grids with real-valued segment lengths is analyzed first, in \S\ref{part: max: arbitrary grid, real lengths}. The particular case of square grids is addressed in \S\ref{part: max: square grid, real lengths}, as it allows a specialized formula for the direct problem. Lastly, the analysis of a unit square grid with integer-valued lengths is given in \S\ref{part: max: unit square grid, integer lengths}.

@@

\section{Average number of tiles}
\label{part: ave}

@@

***Maybe a graph comparing maximum and average. Or compute/plot their ratio. The asymptotic slopes are $\sqrt(1/a^2+1/b^2)$ for $\funt$ and $2/\pi(1/a+1/b)$. It may be interesting to consider the ratio of asymptotic slopes. Some quick plot of that ratio suggests it is maximum for $a=b$, and equals $sqrt(8)/pi = 0.9003$.

>>>>

The problem of how many tiles a \emph{random} segment of a given length touches on average is precisely defined as follows. By symmetry, one endpoint of the segment can be assumed to lie in a fixed, reference tile. The position of this endpoint is \emph{uniformly} distributed on the tile. The segment orientation has a \emph{uniform} distribution on the set of all possible directions, and is \emph{independent} of the endpoint position. As will be seen, solving this problem also answers the inverse question (segment length to touch a given number of tiles on average), which thus need not be considered separately.

@@

The problems studied in this paper are related to \ldots ***maybe a short paragraph here mentioning related problems: Buffon's needle, Buffon-Laplace problem. Some references: ``Buffon's Noodle Problem'', J.F.~Ramaley, 1969. ``On Laplace's Extension of the Buffon Needle Problem'', B.J.~Arnow, 1994. \\
\url{https://www.jstor.org/stable/2317945} \\
\url{https://www.jstor.org/stable/2687085}

The rest of the paper is organized as follows. Fundamental results are presented in \S\ref{part: fund results}, which form the basis of the subsequent analysis. The direct and inverse problems for a deterministic segment are considered in \S\ref{part: max}, first for real-valued lengths, and then for integer-valued lengths on a unit square grid. The problem of computing the average number of tiles touched by a random segment is analyzed in \S\ref{part: ave}.

@@

\section{Direct and inverse problems for a deterministic segment}
\label{part: max}
% ***Is it clear enough that "maximum number of tiles" can refer to the inverse problem too? (because the inverse problem is closely related to the direct one, as discussed in the introduction)
% If not, an alernative would be a more explicit title like: "Maximum number of tiles fo a given length. Conditions on length for a given number of tiles"

The direct and inverse problems defined in \S\ref{part: intro}, considering the segment position and orientation as deterministic, are addressed in this section. The general case for rectangular grids with real-valued segment lengths is analyzed first, in \S\ref{part: max: arbitrary grid, real lengths}. The particular case of square grids is addressed in \S\ref{part: max: square grid, real lengths}, as it allows a specialized formula for the direct problem. Lastly, the analysis of a unit square grid with integer-valued lengths is given in \S\ref{part: max: unit square grid, integer lengths}.

@@

\section{Average number of tiles touched by a random segment}
\label{part: ave}

@@

***Maybe a graph comparing maximum and average. Or compute/plot their ratio. The asymptotic slopes are $\sqrt(1/a^2+1/b^2)$ for $\funt$ and $2/\pi(1/a+1/b)$. It may be interesting to consider the ratio of asymptotic slopes. Some quick plot of that ratio suggests it is maximum for $a=b$, and equals $sqrt(8)/pi = 0.9003$. The minimum number of tile that can be touched with non-zero probability (corresponding to an n times 1 canonical rectangle if a>=b) can also be plotted.
